We begin our empirical analysis by estimating the causal effect of app access on household electricity consumption, addressing our first research question: how persistent are feedback-induced savings from continuous smartphone feedback? Given the staggered rollout across three tranches, we estimate dynamic treatment effects following \citet{sunEstimatingDynamicTreatment2021}:
\[
y_{it} = \sum_{k \neq -1} \beta_k \cdot \mathbb{1}\{t - G_i = k\} + \alpha_i + \delta_t + \varepsilon_{it},
\]
where $y_{it}$ is household $i$'s daily electricity consumption (kWh, aggregated from 15-minute smart meter data), $G_i$ denotes app access timing, $\beta_k$ traces effects relative to the $k=-1$ pre-treatment reference, $\alpha_i$ absorbs time-invariant household differences, and $\delta_t$ controls common shocks (weekly or daily fixed effects). Standard errors are clustered at the household level.

Table \ref{table:main_results} presents the main results. Across specifications, we observe a consistent reduction in electricity consumption following access to the application (columns 1--3). In the baseline specification (column 1), the TWFE estimate indicates a decrease of 0.47 kWh per household per day (SE = 0.206, $p = 0.023$). The corresponding Sun--Abraham estimate is similar in magnitude ($-0.45$ kWh, SE = 0.294, $p = 0.128$), suggesting that accounting for staggered treatment timing and heterogeneous effects does not materially alter the point estimate, although statistical precision is reduced. Including weekly fixed effects yields nearly identical results, with estimated reductions of 0.46--0.47 kWh (column 3). In proportional terms, this corresponds to a reduction of approximately 4.7\% in electricity consumption.\footnote{The percentage effect is based on a log-linear specification using $\log(\text{consumption})$ as the dependent variable.} 

From a behavioral perspective, these results are consistent with a salience-based mechanism: by continuously providing feedback, the application makes electricity consumption more cognitively accessible, thereby increasing the likelihood that households adjust their behavior. The persistence of the effect suggests that this is not merely a short-lived attention shock. 

To further investigate the underlying behavioral adjustments, we decompose total consumption into flexible (hours 7--22) and non-flexible (hours 23--6) components. The reduction in electricity use is primarily driven by flexible consumption, where households have greater scope for behavioral adaptation. In this domain, the estimated treatment effect ranges from 0.31 to 0.33 kWh (e.g., column 4: $-0.31$ kWh, SE = 0.156, $p = 0.048$), indicating economically meaningful reductions. In contrast, the effect on non-flexible consumption is smaller (approximately $-0.14$ to $-0.15$ kWh; column 5: $p = 0.082$) and only weakly statistically significant.



\begin{table}[htbp]\centering
\caption{Treatment Effects on Electricity Consumption}
\label{table:main_results}
\begin{threeparttable}
\begin{tabular}{lccccc}
\toprule
 & 1 & 2 & 3 & 4 & 5 \\
\midrule

\textbf{TWFE} \\
Treatment effect 
& -0.4685** 
& -0.0187** 
& -0.4612** 
& -0.3107** 
& -0.1526* \\
& (0.2061) 
& (0.0086) 
& (0.2059) 
& (0.1562) 
& (0.0784) \\

\addlinespace
\textbf{Sun--Abraham ATT} \\
ATT 
& -0.4500 
& --- 
& -0.4736** 
& -0.3323* 
& -0.1424 \\
& (0.2938) 
& --- 
& (0.2343) 
& (0.1730) 
& (0.0939) \\

\midrule
Observations 
& 354,121 
& 8,287,525 
& 354,121 
& 354,096 
& 354,121 \\

Household FE & Yes & Yes & Yes & Yes & Yes \\
Time FE & Day & Day & Week & Week & Week \\

\bottomrule
\end{tabular}

\begin{tablenotes}
\footnotesize
\item Notes: The table reports treatment effects on electricity consumption (kWh). 
TWFE estimates are obtained from two-way fixed effects regressions. Sun--Abraham estimates report the aggregated ATT from staggered difference-in-differences models. 
All standard errors are clustered at the household level. 
Significance levels: ** $p<0.05$, * $p<0.10$.
\end{tablenotes}

\end{threeparttable}
\end{table}